---
title: "Initial Analysis of Police Data"
author: "George Pope, 300416638"
date: "24/08/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
PS <- read.csv("Proceedings_Police_Stations.csv")
OD <- read.csv("Proceedings_Offender_Demographics.csv")

library(ggplot2)
library(dplyr)
library(tidyr)
library(data.table)
```

Crimes committed by Europeans and Maori:
```{r}
ethnicityCrime <- OD[ , c(2, 9)]
names(ethnicityCrime)[names(ethnicityCrime)=="Anzsoc.Division"] <- "Crime.Type"

countCrime <- dcast(setDT(ethnicityCrime, keep.rownames = TRUE), Ethnicity ~ Crime.Type, length)
names(countCrime)[names(countCrime)=="Anzsoc.Division"] <- "Crime.Type"
names(countCrime) <- c("Ethnicity", 'Abduction/Harassment', "Assault", "Endagerment", "Fraud", "Homicide",
                       "Drugs", "Miscellaneous", "Govt/Justice", "Weapons", "Property/Environment", "Public Order",
                       "Robbery/Extortion", "Sexual Assault", "Theft", "Traffic/Vehicle", "Break and Enter")

#Crimes committed by Europeans and Maori

eu.maCrime <- countCrime[c(3,6)]
eu.maCrime <- as.data.frame(t(eu.maCrime))

names(eu.maCrime)[1] <- "European"
names(eu.maCrime)[2] <- "Maori"
eu.maCrime <- eu.maCrime[-1,]
eu.maCrime

#Changing index

eu.maCrime$Type <- rownames(eu.maCrime)
eu.maCrime$Index <- 1:nrow(eu.maCrime)
row.names(eu.maCrime) <- eu.maCrime$Index
eu.maCrime$Index <- NULL

#Pivoting data

eu.maCrime <- eu.maCrime %>% select(Type, Maori, European) %>% 
  pivot_longer(cols = c(Maori, European), names_to = "Var", values_to = "Val")

#Changing datatype of Val from character to integer

eu.maCrime$Val <- as.integer(eu.maCrime$Val)

#Plot

ggplot(data = eu.maCrime, mapping = aes(x=Type, y=Val, fill=Var)) +
  geom_bar(stat = "identity", position=position_dodge()) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Types of Crimes Committed by Maori and Europeans", fill = "Ethnicity") + 
  xlab("Type") + ylab("Frequency")
```
The variables along the x-axis refer to the types of crimes committed by particular ethnicities. The naming of types have been simplified to allow easy fit onto the graph.The y-axis refers to the number of times these types of crimes occurred. 

It is interesting that despite the Maori population being far smaller than the European population in New Zealand, they have similar crime rates, even exceeding Europeans in certain crime types, such as theft and assault. 


Types of crimes committed in certain regions:
```{r}
#Types of crimes committed per region

crimeLocation <- OD[ ,c(2,17)]
names(crimeLocation)[1] <- "Crime.Type"
names(crimeLocation)[2] <- "Location"

countCrime.location <- dcast(setDT(crimeLocation, keep.rownames = TRUE), Location ~ Crime.Type, length)
names(countCrime.location) <- c("Location", 'Abduction/Harassment', "Assault", "Endagerment", "Fraud", "Homicide",
                       "Drugs", "Miscellaneous", "Govt/Justice", "Weapons", "Property/Environment", "Public Order",
                       "Robbery/Extortion", "Sexual Assault", "Theft", "Traffic/Vehicle", "Break and Enter")

#Crimes for Aukland and Wellington

countCrime.location <- countCrime.location[c(1,14), ]
countCrime.location <- as.data.frame(t(countCrime.location))
names(countCrime.location)[1] <- "Auckland"
names(countCrime.location)[2] <- "Wellington"
countCrime.location <- countCrime.location[-1,]
countCrime.location

#Changing Index

countCrime.location$Type <- rownames(countCrime.location)
countCrime.location$Index <- 1:nrow(countCrime.location)
row.names(countCrime.location) <- countCrime.location$Index
countCrime.location$Index <- NULL

#Pivoting data

countCrime.location <- countCrime.location %>% select(Type, Auckland, Wellington) %>% 
  pivot_longer(cols = c(Auckland, Wellington), names_to = "Var", values_to = "Val")

#Changing datatype of Val from character to integer

countCrime.location$Val <- as.integer(countCrime.location$Val)

#Plot

ggplot(data = countCrime.location, mapping = aes(x=Type, y=Val, fill=Var)) +
  geom_bar(stat = "identity", position=position_dodge()) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Types of Crimes Committed in the Auckland City and Wellington Districts", fill = "District") + 
  xlab("Type") + ylab("Frequency")
```
It is important to note that Auckland here refers to the Auckland City district, not the entirety of Auckland. Wellington in this instance refers to the entirety of the Wellington district, which includes areas like Upper Hutt. 

Each of the variables along the x-axis are types of crimes that are committed. The naming of types have been simplified to allow easy fit onto the graph. The y-axis is the number of times specific crimes have been committed in the district. 
It is interesting to note that although the Auckland City District has the greatest population out of all possible districts, it has a very similar number of crimes compared to the Wellington District.


Age group of offenders

```{r}
#Age group of offenders

crimeAge <- PS[ ,c(2,6)]
names(crimeAge)[1] <- "Crime.Type"
names(crimeAge)[2] <- "Age"

countCrime.Age <- dcast(setDT(crimeAge, keep.rownames = TRUE), Age ~ Crime.Type, length)
names(countCrime.Age)[names(countCrime)=="Anzsoc.Division"] <- "Crime.Type"

countCrime.Age$Total <- rowSums(countCrime.Age[,2:17])
countCrime.Age

countCrime.Age <- countCrime.Age[ , c(1, 18)]
countCrime.Age <- countCrime.Age[c(2,3,4,5,7,8,9,11,12,13,14,15,16,17), ]

ggplot(data = countCrime.Age, mapping = aes(x=Age, y=Total)) +
  geom_bar(stat = "identity", fill = "#3366CC") +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Total Crimes Committed by Age Group") + 
  xlab("Age") + ylab("Frequency")
```
Note ages, 0-4, 5-9, NotApplicable and NotSpecified were removed as they were so small compared to the other age groups.

The x-axis refers to the age groups of offenders and the y-axis is the total number of crimes recorded for that specific age group.

A point of interest is that crimes occur the most commonly at ages 15 through to 29, and then begins to taper off as you get older. There is a steady decline in crimes commited as you reach age 30, at a decreasing rate. 


Crimes committed by youths (15-24) and people aged 40-54

```{r}
#Types of crimes for youths i.e. 15-24

crimeAge <- PS[ ,c(2,6)]
names(crimeAge)[1] <- "Crime.Type"
names(crimeAge)[2] <- "Age"

countCrime.Age <- dcast(setDT(crimeAge, keep.rownames = TRUE), Age ~ Crime.Type, length)
names(countCrime.Age)[names(countCrime)=="Anzsoc.Division"] <- "Crime.Type"
names(countCrime.Age) <- c("Age", 'Abduction/Harassment', "Assault", "Endagerment", "Fraud", "Homicide",
                                "Drugs", "Miscellaneous", "Govt/Justice", "Weapons", "Property/Environment", "Public Order",
                                "Robbery/Extortion", "Sexual Assault", "Theft", "Traffic/Vehicle", "Break and Enter")

countCrime.Age <- countCrime.Age[c(3,4), ]
countCrime.Age <- as.data.frame(t(countCrime.Age))

#Crimes for specific ages

names(countCrime.Age)[1] <- "15-19"
names(countCrime.Age)[2] <- "20-24"
countCrime.Age <- countCrime.Age[-1,]

#Changing Index

countCrime.Age$Type <- rownames(countCrime.Age)
countCrime.Age$Index <- 1:nrow(countCrime.Age)
row.names(countCrime.Age) <- countCrime.Age$Index
countCrime.Age$Index <- NULL

countCrime.Age$"15-19" <- as.integer(countCrime.Age$"15-19")
countCrime.Age$"20-24" <- as.integer(countCrime.Age$"20-24")

countCrime.Age$Total <- rowSums(countCrime.Age[,c(1,2)])
countCrime.Age <- countCrime.Age[ ,c(3,4)]

ggplot(data = countCrime.Age, mapping = aes(x=Type, y=Total)) +
  geom_bar(stat = "identity", fill = "#3366CC") +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Crimes Committed by Youths (15-24)") + 
  xlab("Type of Crime") + ylab("Frequency")

#----------------------------------------------------------------------------------------

#Crimes for older age

crimeOldAge <- PS[ ,c(2,6)]
names(crimeOldAge)[1] <- "Crime.Type"
names(crimeOldAge)[2] <- "Age"

countCrime.OldAge <- dcast(setDT(crimeOldAge, keep.rownames = TRUE), Age ~ Crime.Type, length)
names(countCrime.OldAge)[names(countCrime)=="Anzsoc.Division"] <- "Crime.Type"
names(countCrime.OldAge) <- c("Age", 'Abduction/Harassment', "Assault", "Endagerment", "Fraud", "Homicide",
                           "Drugs", "Miscellaneous", "Govt/Justice", "Weapons", "Property/Environment", 
                           "Public Order","Robbery/Extortion", "Sexual Assault", "Theft", "Traffic/Vehicle", 
                           "Break and Enter")

countCrime.OldAge <- countCrime.OldAge[c(8,9,10,11), ]
countCrime.OldAge <- as.data.frame(t(countCrime.OldAge))

#Crimes for specific ages

names(countCrime.OldAge)[1] <- "40-44"
names(countCrime.OldAge)[2] <- "45-49"
names(countCrime.OldAge)[3] <- "50-54"
names(countCrime.OldAge)[4] <- "55-59"
countCrime.OldAge <- countCrime.OldAge[-1,]

#Changing Index

countCrime.OldAge$Type <- rownames(countCrime.OldAge)
countCrime.OldAge$Index <- 1:nrow(countCrime.OldAge)
row.names(countCrime.OldAge) <- countCrime.OldAge$Index
countCrime.OldAge$Index <- NULL

countCrime.OldAge$"40-44" <- as.integer(countCrime.OldAge$"40-44")
countCrime.OldAge$"45-49" <- as.integer(countCrime.OldAge$"45-49")
countCrime.OldAge$"50-54" <- as.integer(countCrime.OldAge$"50-54")
countCrime.OldAge$"55-59" <- as.integer(countCrime.OldAge$"55-59")

countCrime.OldAge$Total <- rowSums(countCrime.OldAge[,c(1,2,3,4)])
countCrime.OldAge <- countCrime.OldAge[ ,c(5,6)]

ggplot(data = countCrime.OldAge, mapping = aes(x=Type, y=Total)) +
  geom_bar(stat = "identity", fill = "#CC3300") +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Crimes Committed by people aged 40-54") + 
  xlab("Type of Crime") + ylab("Frequency")

#--------------------------------------------------------------------------------------------

#Comparison of age groups via stacked bar chart

#Renaming total columns

youngCrime <- countCrime.Age
oldCrime <- countCrime.OldAge
names(youngCrime)[2] <- "Youth"
names(oldCrime)[2] <- "Old"

#Combining young and old statistics

compareCrime <- cbind(youngCrime, oldCrime)
compareCrime[3] <- NULL

#Pivot to enable plotting on ggplot

compareCrime.edit <- compareCrime %>% select(Type, Youth, Old) %>% 
  pivot_longer(cols = c(Youth, Old), names_to = "Var", values_to = "Val")

ggplot(data = compareCrime.edit, mapping = aes(x=Type, y=Val, fill = Var)) +
  geom_bar(stat = "identity", position = "fill") +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Proportion of Crimes Committed by Youths (15-24) and Old (40-59)", fill = "Age group") + 
  xlab("Type of Crime") + ylab("Proportion")
```
Each of the variables along the x-axis are types of crimes that are committed. The naming of types have been simplified to allow easy fit onto the graph. The y-axis is the number of times specific crimes have been committed by particular age brackets.

The most common types of crime for youths to commit are assault, theft, and traffic/vehicle offences, which are similar to the most common crimes committed by people in the age bracket 40-54, with the exception of Offences Against Justice Procedures, Govt Sec and Govt Ops (Govt/Justice). Theft is also much more common for youths when compared to the older age group.

Assault is the most common crime for the older age bracket, whereas traffic/vehicle offences are the most common for youths. 

```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```