---
title: "Recorded Crime Victim Statistics (RCVS), Police Stations"
subtitle: "(Datasets Policedata.nz: AEG_Full_Data_data.csv, UAEG_Full_Data_data.csv)"
author: "EDA, Maria DaRocha"
date: "8/30/2020"
output: pdf_document
---

```{r,echo=FALSE,message=FALSE,warning=FALSE}
require(knitr)
require(formatR)
# Set so that long lines in R will be wrapped:
opts_chunk$set(tidy.opts=list(width.cutoff=40),tidy=TRUE)
opts_knit$set(global.par = TRUE)
## Reduce the amount of whitespace above each plot
if(is_latex_output()) {
  plot_default <- knit_hooks$get("plot")
  knit_hooks$set(plot = function(x, options) { 
    x <- c("\\vspace{-2mm}", plot_default(x, options))
  })
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
VD <- read.csv("AEG_Full_Data_data.csv")
UVD <- read.csv("UAEG_Full_Data_data.csv")

library(ggplot2)
library(dplyr)
library(tidyr)
library(data.table)
library(RColorBrewer)
```
  
# Victimisations (Police Stations): 
## Crime Division, Ethnicities of Victims   
  
```{r}
division_ethnicity <- VD[ , c(2, 9)]
names(division_ethnicity)[names(division_ethnicity)=="Anzsoc.Division"] <- "Division"
names(division_ethnicity)[names(division_ethnicity)=="Abduction, Harassment and Other Related Offences Against a Person"] <- "Abduction"

victimisation_count <- dcast(setDT(division_ethnicity, keep.rownames = TRUE), Ethnicity ~ Division, length)

kable(head(victimisation_count))
```


```{r}
# Visual Observation: victimisation ONE crime division with Bar Chart
ggplot(data = victimisation_count, mapping = aes(x=victimisation_count$Ethnicity, y=victimisation_count$`Abduction`, fill=victimisation_count$Ethnicity)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle =70, hjust=1)) +
  labs(title = "Abduction and Harassment, Victimisations", subtitle = '*Includes Related Offences (2015-2020)', fill = "Ethnicity") + xlab("Ethnicity") + ylab("Recorded Victimisations")
```
  
# Victimisations (Police Stations): 
## Crime Division, (Aggregated) Ethnicities of Victims   

```{r}
# Aggregated Minority (Other)
# Retain 'Not Stated'
aggregation <- c(0,0,1,0,0,2,0,0,0,3,0,0,0)
victimisation_count_copy <- victimisation_count
victimisation_count_copy$Ethnicity <- aggregation
aggregate_victimisation <- aggregate(x=victimisation_count_copy, by=list(victimisation_count_copy$Ethnicity), FUN = sum)
names(aggregate_victimisation)[names(aggregate_victimisation)=="Abduction, Harassment and Other Related Offences Against a Person"] <- "Abduction"
aggregate_victimisation$Ethnicity <- c("Minority (Other)", "European", "Maori", "Not Stated")
# drop added grouping
aggregate_victimisation <- aggregate_victimisation[,-(1)]
kable(head(aggregate_victimisation))
```


```{r}
# Visual Observation of ONE division with Bar Chart
ggplot(data = aggregate_victimisation, mapping = aes(x=Ethnicity, y=`Abduction`, fill=Ethnicity)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 0)) +
  labs(title = "Abduction and Harassment, Victimisations", subtitle = '*Includes Related Offences (2015-2020)', fill = "Ethnicity") + xlab("Ethnicity") + ylab("Recorded Victimisations")
```
  
# Victimisation (Police Stations): 
## Crime Division to Year of Report  

```{r}
victimisation_series <- VD[ , c(2, 5)]

names(victimisation_series)[names(victimisation_series)=="Anzsoc.Division"] <- "Division"
series_count <- dcast(setDT(victimisation_series, keep.rownames = TRUE), Year.Month ~ Division, length)

```

```{r}
agg_series_count <- series_count
years <- c(2015, 2016, 2017, 2018, 2019, 2020)
for (k in years){
  year <- grepl(k, agg_series_count$Year.Month)
  index <- 0
  for (y in year){
    index <- index + 1
    if (y==TRUE){
      agg_series_count$Year.Month[index] = k
    }
  }
}  

names(agg_series_count)[names(agg_series_count)=="Year.Month"] <- "Year"
names(agg_series_count)[names(agg_series_count)=="Abduction, Harassment and Other Related Offences Against a Person"] <- "Abduction"

agg_series_count <- aggregate(cbind(`Abduction`, `Acts Intended to Cause Injury`, `Robbery, Extortion and Related Offences`, `Sexual Assault and Related Offences`, `Theft and Related Offences`, `Unlawful Entry With Intent/Burglary, Break and Enter`) ~ Year, agg_series_count, sum)

kable(head(agg_series_count))
```


```{r}
# Set up visualisation
all_colnames <- c("Date","Abduction","Intended_Injury","Robbery","Sexual_Assault","Theft","Burglary")
index <- 0
for (col in all_colnames){
  index <- index + 1
  names(agg_series_count)[index] <- all_colnames[index]
}

agg_series_count <- agg_series_count %>%
  select(Date,Abduction,Intended_Injury,Robbery,Sexual_Assault,Theft,Burglary) %>%
  gather(key = variable, value = value, -Date)

```
  
# Time Series, Victimisations (Police Stations):
## Change in Recorded Victimisations by Crime Division (Annually) 
  
```{r}
# Draw Time Series
ggplot(agg_series_count, aes(x = Date, y = value, group = variable)) + 
  geom_line(aes(color = variable, linetype = variable)) + 
  scale_color_brewer(palette = "Dark2") +
  theme(axis.text.x=element_text(angle = 90, size = 6, hjust=1)) +
  labs(title = "Victimisations by Crime Division by Year", subtitle = 'Time Series Data (2015-2020)', fill = "Crime Division") + xlab("Year") + ylab("Recorded Victimisations")

```

# Unique Victims (Police Stations): 
## Crime Division, (Aggregated) Ethnicities of Victims  
  

```{r}
#head(UVD)
uv_count <- UVD[ , c(2, 8)]
names(uv_count)[names(uv_count)=="Anzsoc.Division"] <- "Division"

uv_count <- dcast(setDT(uv_count, keep.rownames = TRUE), Ethnicity ~ Division, length)

# Aggregated Minority (Other)
# Retain 'Not Stated'
aggregation <- c(0,0,1,0,0,2,0,0,0,3,0,0,0)
uv_count_copy <- uv_count
uv_count_copy$Ethnicity <- aggregation
aggregate_uv <- aggregate(x=uv_count_copy, by=list(uv_count_copy$Ethnicity), FUN = sum)
names(aggregate_uv)[names(aggregate_uv)=="Abduction, Harassment and Other Related Offences Against a Person"] <- "Abduction"
aggregate_uv$Ethnicity <- c("Minority (Other)", "European", "Maori", "Not Stated")
# drop added grouping
aggregate_uv <- aggregate_uv[,-(1)]

kable(head(aggregate_uv))
```
```{r}
# Visual Observation of ONE division with Bar Chart
ggplot(data = aggregate_uv, mapping = aes(x=Ethnicity, y=`Abduction`, fill=Ethnicity)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 0)) +
  labs(title = "Abduction and Harassment, Unique Victims", subtitle = '*Includes Related Offences (2015-2020)', fill = "Ethnicity") + xlab("Ethnicity") + ylab("Recorded Victimisations")

```

# Unique Victims & Victimisations (Police Stations):
## Distribution Analysis with Standard Deviation

```{r}
distribution_names <- colnames(aggregate_uv[1,])

distribution_names <- distribution_names[-1]


# Comparing Distributions
set.seed(1234)
for (n in 1:length(distribution_names)){
  name <- distribution_names[n]
  
  uv_vals <- aggregate_uv[ , grepl(name, names(aggregate_uv))]
  v_vals <- aggregate_victimisation[ , grepl(name, names(aggregate_uv))]
  
  uv_group <- data.frame(x = aggregate_uv$Ethnicity, 
                  y = uv_vals)
  uv_group$group=c(0,0,0,0)
  uv_group <- uv_group %>% mutate_at(c("y"), ~(scale(.) %>% as.vector))
  
  v_group <- data.frame(x = aggregate_victimisation$Ethnicity, 
                    y = v_vals)
  v_group$group=c(1,1,1,1)
  v_group <- v_group %>% mutate_at(c("y"), ~(scale(.) %>% as.vector))
  
  # Merge dataframes
  distribution_analysis <- merge(uv_group, v_group, by.x="x", by.y="x")

  # Pivot data
  distribution_analysis_copy <- distribution_analysis %>% select(x, y.x,y.y) %>% 
  pivot_longer(cols = c(y.x,y.y), names_to = "variable", values_to = "value")
  # print(distribution_analysis_copy)
  
  # Plot
  p <- ggplot(distribution_analysis_copy, aes(x = x, y = value, group = variable)) + 
  geom_line(aes(color = variable, linetype = variable)) + 
  scale_color_brewer(palette = "Dark2") +
  theme(axis.text.x=element_text(angle = 0, size = 10), legend.position = "none") +
  labs(title = name, 
       subtitle = 'Fluxuation in Distribution of Ethnicities: \nUnique Victims and Victimisations (2015-2020)', fill = name) + xlab("Ethnicity") + ylab("Stnd Deviation (Mean)")
  show(p)
}

```




