---
title: "Police Dashboard: Exploratory Data Analysis (EDA)"
subtitle: "DATA301: Group Project"
author: "Maria DaRocha (300399718), George Pope (300416638), Usanthini Sasikumar (300451063)"
date: "8/30/2020"
output: pdf_document
---

```{r,echo=FALSE,message=FALSE,warning=FALSE}
require(knitr)
require(formatR)
# Set so that long lines in R will be wrapped:
opts_chunk$set(tidy.opts=list(width.cutoff=40),tidy=TRUE)
opts_knit$set(global.par = TRUE)
## Reduce the amount of whitespace above each plot
if(is_latex_output()) {
  plot_default <- knit_hooks$get("plot")
  knit_hooks$set(plot = function(x, options) { 
    x <- c("\\vspace{-2mm}", plot_default(x, options))
  })
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# GitHub Repository:  
https://github.com/Marianette/Police-Dashboard

The datasets that were used for this report are Offender Proceedings (offender demographics), Offender Proceedings (police stations), Police Dataset (Victimisation Time and Place), Victimization (Police Stations) and Unique Victims (Police Stations).  

## Background and Description of Datasets Offender Proceedings (Offender Demographics) and Offender Proceedings (Police Stations)  

These datasets are of particular interest as they show each crime committed, the type of crime, age, gender, the ethnicity of the offender, and where it was committed. It could be used to identify which types of crime are most common in certain areas of New Zealand, and which demographic is most likely to commit these of crimes.  

The types of data present in these datasets are both numerical and categorical. The types of offenses, location, and other variables are stored as characters, as well as time-series data, and integers for certain variables. There are some values missing in the dataset, where they were not stated, or where people’s information had been suppressed. 

## Background and Description of Datasets Police Dataset (Victimisation Time and Place)  

This dataset has 107089 observations of 14 variables. This dataset is of particular interest as it shows each crime committed, the type of crime, when it was committed, and where it was committed. It could be used to see and identify if there are any links or patterns between the type of crime committed and the area of the crime committed (Territorial authority) around Wellington. 

The types of data present in this dataset are both numerical and categorical. The types of crimes, location, and other variables are stored as characters, as well as time-series data and Integers for certain variables such as the number of records, victimizations, and other variables. There are some values missing in the dataset, where they were not listed clearly, or where the information had been reduced. 


## Background and Description of Datasets Victimisations & Unique Victims (Police Stations)

Among the datasets we decided to work on, two of them were: Victimisations (Police Stations), and Unique Victims (Police Stations). These datasets were of particular interest because oftentimes when we discuss crime, the majority of our conversations revolve around the offender: their background, their demographic, etcetera. While these details may help us better understand an individual’s propensity towards crime, focusing solely on the offender can mean overlooking the other involved party: their victims. Thoughtfully done, analyzing information about victims can lead to a greater understanding of commonalities between victims of similar crimes. This layer of insight could help us better understand what might make someone more vulnerable to crime, or susceptible to being targeted by a specific type of crime. 

The datasets themselves consisted of largely *categorical data*. The data is best suited to time series models, as it is intended to offer insights about victimisations and unique victims within a specific time period (2015-2020).  

While there were no missing values in the Victimisations (Police Stations) dataset, there were some corrupted entries in which a categorical variable (Ages 10-19) had become stored as a date (Oct-19). This issue could be related to the files being exported from policedata.nz as ‘Microsoft Excel csv files,’ instead of regular csv files. Data integration took place on an as-needed basis. For example, when comparing ethnicity and crime division features in each dataset, the four features were merged into a single dataframe. The Unique Victims dataset was missing just over 68,000 values.


```{r, echo=FALSE}
VD <- read.csv("AEG_Full_Data_data.csv")
UVD <- read.csv("UAEG_Full_Data_data.csv")

library(ggplot2)
library(dplyr)
library(tidyr)
library(data.table)
library(RColorBrewer)
```

# General EDA, Victimisations (Police Stations): 
## Missing Values, Data Types, Etc.

```{r, echo=FALSE}
# Trivial investigation
summary(VD)
test_NA_VD <- sum(is.na(VD))
test_NA_UVD <- sum(is.na(UVD))

test_NA_UVD
test_NA_VD

# Not expressive of corrupt data (date values in place of age values, VD)
```

  
# Victimisations (Police Stations): 
## Crime Division, Ethnicities of Victims   

In the victimisations (by police stations) dataset, we chose to explore ethnicity’s connection to each crime division. If patterns are found in later, deeper analysis - then we will be able to conclude some correlation between ethnicity and a particular type of crime. In a real-world application, this would help identify a population that may be more vulnerable to certain crimes. Initially, we explored the data using the classifications for ethnicity provided. Below, we show the distributions of Ethnicity classes for crimes involving abduction and sexual assault.
  
```{r, echo=FALSE}
division_ethnicity <- VD[ , c(2, 9)]
names(division_ethnicity)[names(division_ethnicity)=="Anzsoc.Division"] <- "Division"
names(division_ethnicity)[names(division_ethnicity)=="Abduction, Harassment and Other Related Offences Against a Person"] <- "Abduction"

victimisation_count <- dcast(setDT(division_ethnicity, keep.rownames = TRUE), Ethnicity ~ Division, length)

kable(head(victimisation_count))
```


```{r, echo=FALSE}
# Visual Observation: victimisation ONE crime division with Bar Chart
ggplot(data = victimisation_count, mapping = aes(x=victimisation_count$Ethnicity, y=victimisation_count$`Abduction`, fill=victimisation_count$Ethnicity)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle =70, hjust=1)) +
  labs(title = "Abduction and Harassment, Victimisations", subtitle = '*Includes Related Offences (2015-2020)', fill = "Ethnicity") + xlab("Ethnicity") + ylab("Recorded Victimisations")
```
```{r, echo=FALSE}
# Visual Observation: victimisation ONE crime division with Bar Chart
ggplot(data = victimisation_count, mapping = aes(x=victimisation_count$Ethnicity, y=victimisation_count$`Sexual Assault and Related Offences`, fill=victimisation_count$Ethnicity)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle =70, hjust=1)) +
  labs(title = "Sexual Assault, Victimisations", subtitle = '*Includes Related Offences (2015-2020)', fill = "Ethnicity") + xlab("Ethnicity") + ylab("Recorded Victimisations")
```
  
# Victimisations (Police Stations): 
## Crime Division, (Aggregated) Ethnicities of Victims   

Upon seeing the data heavily reflect European, Maori, and unspecified (or suppressed) ethnicities, it seemed appropriate to aggregate non-Maori and non-Eurpoean ethnic groups into a "Minority (Other)" category. This aggregation provides a visual model which improves the comparability of minority groups that are less prevalent New Zealand. In doing this, we ensure that individuals who separately fall into the (now aggregated) "Minority (other)" category are fairly represented - and that visually, the model no longer reflects any one minority category appearing less targeted than Maori or European persons.

```{r, echo=FALSE}
# Aggregated Minority (Other)
# Retain 'Not Stated', 'Not Applicable', 'Not Elsewhere Classified'
# 'Organisation', 'Suppressed'
aggregation <- c(0,0,1,0,0,2,0,3,4,5,6,0,7)
victimisation_count_copy <- victimisation_count
victimisation_count_copy$Ethnicity <- aggregation
aggregate_victimisation <- aggregate(x=victimisation_count_copy, by=list(victimisation_count_copy$Ethnicity), FUN = sum)
names(aggregate_victimisation)[names(aggregate_victimisation)=="Abduction, Harassment and Other Related Offences Against a Person"] <- "Abduction"
aggregate_victimisation$Ethnicity <- c("Minority (Other)", "European", "Maori", "Not Applicable", "Not Elsewhere Classified", "Not Stated", "Organisation", "Suppressed")
# drop added grouping
aggregate_victimisation <- aggregate_victimisation[,-(1)]
kable(head(aggregate_victimisation))
```


```{r, echo=FALSE}
# Visual Observation of ONE division with Bar Chart
ggplot(data = aggregate_victimisation, mapping = aes(x=Ethnicity, y=`Abduction`, fill=Ethnicity)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 75, hjust=1)) +
  labs(title = "Abduction and Harassment, Victimisations", subtitle = '*Includes Related Offences (2015-2020)', fill = "Ethnicity") + xlab("Ethnicity") + ylab("Recorded Victimisations")
```
  
# Victimisation (Police Stations): 
## Crime Division to Year of Report  

The next interesting feature of the data is "Year.Month" - i.e. the basis for our time series. In future analysis, we will investigate this feature in greater detail. For now, we can aggregate these values to just the year on record and observe basic trends.

```{r, echo=FALSE}
victimisation_series <- VD[ , c(2, 5)]

names(victimisation_series)[names(victimisation_series)=="Anzsoc.Division"] <- "Division"
series_count <- dcast(setDT(victimisation_series, keep.rownames = TRUE), Year.Month ~ Division, length)

```

```{r, echo=FALSE}
agg_series_count <- series_count
years <- c(2015, 2016, 2017, 2018, 2019, 2020)
for (k in years){
  year <- grepl(k, agg_series_count$Year.Month)
  index <- 0
  for (y in year){
    index <- index + 1
    if (y==TRUE){
      agg_series_count$Year.Month[index] = k
    }
  }
}  

names(agg_series_count)[names(agg_series_count)=="Year.Month"] <- "Year"
names(agg_series_count)[names(agg_series_count)=="Abduction, Harassment and Other Related Offences Against a Person"] <- "Abduction"

agg_series_count <- aggregate(cbind(`Abduction`, `Acts Intended to Cause Injury`, `Robbery, Extortion and Related Offences`, `Sexual Assault and Related Offences`, `Theft and Related Offences`, `Unlawful Entry With Intent/Burglary, Break and Enter`) ~ Year, agg_series_count, sum)

kable(head(agg_series_count))
```


```{r, echo=FALSE}
# Set up visualisation
all_colnames <- c("Date","Abduction","Intended_Injury","Robbery","Sexual_Assault","Theft","Burglary")
index <- 0
for (col in all_colnames){
  index <- index + 1
  names(agg_series_count)[index] <- all_colnames[index]
}

agg_series_count <- agg_series_count %>%
  select(Date,Abduction,Intended_Injury,Robbery,Sexual_Assault,Theft,Burglary) %>%
  gather(key = variable, value = value, -Date)

```
  
# Time Series, Victimisations (Police Stations):
## Change in Recorded Victimisations by Crime Division (Annually) 

Below, we model this abstracted data.
  
```{r, echo=FALSE}
# Draw Time Series
ggplot(agg_series_count, aes(x = Date, y = value, group = variable)) + 
  geom_line(aes(color = variable, linetype = variable)) + 
  scale_color_brewer(palette = "Dark2") +
  theme(axis.text.x=element_text(angle = 90, size = 6, hjust=1)) +
  labs(title = "Victimisations by Crime Division by Year", subtitle = 'Time Series Data (2015-2020)', fill = "Crime Division") + xlab("Year") + ylab("Recorded Victimisations")

```

# Unique Victims (Police Stations): 
## Crime Division, (Aggregated) Ethnicities of Victims  

We conduct a near-identical analysis of Unique Victims (Police Stations) below.  

```{r, echo=FALSE}
#head(UVD)
uv_count <- UVD[ , c(2, 8)]
names(uv_count)[names(uv_count)=="Anzsoc.Division"] <- "Division"

uv_count <- dcast(setDT(uv_count, keep.rownames = TRUE), Ethnicity ~ Division, length)

# Aggregated Minority (Other)
# Retain 'Not Stated'
aggregation <- c(0,0,1,0,0,2,0,3,4,5,6,0,7)
uv_count_copy <- uv_count
uv_count_copy$Ethnicity <- aggregation
aggregate_uv <- aggregate(x=uv_count_copy, by=list(uv_count_copy$Ethnicity), FUN = sum)
names(aggregate_uv)[names(aggregate_uv)=="Abduction, Harassment and Other Related Offences Against a Person"] <- "Abduction"
aggregate_uv$Ethnicity <- c("Minority (Other)", "European", "Maori", "Not Applicable", "Not Elsewhere Classified", "Not Stated", "Organisation", "Suppressed")
# drop added grouping
aggregate_uv <- aggregate_uv[,-(1)]

kable(head(aggregate_uv))
```
```{r, echo=FALSE}
# Visual Observation of ONE division with Bar Chart
ggplot(data = aggregate_uv, mapping = aes(x=Ethnicity, y=`Abduction`, fill=Ethnicity)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 75, hjust=1)) +
  labs(title = "Abduction and Harassment, Unique Victims", subtitle = '*Includes Related Offences (2015-2020)', fill = "Ethnicity") + xlab("Ethnicity") + ylab("Recorded Victimisations")

```

# Unique Victims & Victimisations (Police Stations):
## Distributional Analysis with Standard Deviation

For the final part of the exploratory analysis on this data, it seemed like a reasonable inquiry to test the comparablity of the victimisations and unique victims datasets (in regards to variability). Ethnicity was used as a dummy variable to track the consistency of a single feature's variation across the two datasets. We would expect, if the datasets are comparable, for the standard deviations to exhibit the same (or relatively similar trends). This proved to be the case - and going forward, it will be possible to use this test holding other features of interest constant, and potentially use the two datasets interchangeably. **This is invaluable knowledge for avoiding temporally and computationally expensive (unnecessarily redundant) investigations.**

```{r, echo=FALSE}
distribution_names <- colnames(aggregate_uv[1,])

distribution_names <- distribution_names[-1]


# Comparing Distributions
set.seed(1234)
for (n in 1:length(distribution_names)){
  name <- distribution_names[n]
  
  uv_vals <- aggregate_uv[ , grepl(name, names(aggregate_uv))]
  v_vals <- aggregate_victimisation[ , grepl(name, names(aggregate_uv))]
  
  uv_group <- data.frame(x = aggregate_uv$Ethnicity, 
                  y = uv_vals)
  uv_group$group=c(0,0,0,0)
  uv_group <- uv_group %>% mutate_at(c("y"), ~(scale(.) %>% as.vector))
  
  v_group <- data.frame(x = aggregate_victimisation$Ethnicity, 
                    y = v_vals)
  v_group$group=c(1,1,1,1)
  v_group <- v_group %>% mutate_at(c("y"), ~(scale(.) %>% as.vector))
  
  # Merge dataframes
  distribution_analysis <- merge(uv_group, v_group, by.x="x", by.y="x")

  # Pivot data
  distribution_analysis_copy <- distribution_analysis %>% select(x, y.x,y.y) %>% 
  pivot_longer(cols = c(y.x,y.y), names_to = "variable", values_to = "value")
  # print(distribution_analysis_copy)
  
  # Plot
  p <- ggplot(distribution_analysis_copy, aes(x = x, y = value, group = variable)) + 
  geom_line(aes(color = variable, linetype = variable)) + 
  scale_color_brewer(palette = "Dark2") +
  theme(axis.text.x=element_text(angle = 75, size = 10, hjust=1), legend.position = "none") +
  labs(title = name, 
       subtitle = 'Fluxuation in Distribution of Ethnicities:
       \nUnique Victims and Victimisations (2015-2020)', fill = name) + xlab("Ethnicity") + ylab("Stnd Deviation (Mean)")
  show(p)
}

```

# Proceedings (Police Stations) & Proceedings (Offender Demographics):


```{r echo=FALSE}
PS <- read.csv("Proceedings_Police_Stations.csv")
OD <- read.csv("Proceedings_Offender_Demographics.csv")

```

Crimes committed by Europeans and Maori:
```{r echo=FALSE}
ethnicityCrime <- OD[ , c(2, 9)]
names(ethnicityCrime)[names(ethnicityCrime)=="Anzsoc.Division"] <- "Crime.Type"

countCrime <- dcast(setDT(ethnicityCrime, keep.rownames = TRUE), Ethnicity ~ Crime.Type, length)
names(countCrime)[names(countCrime)=="Anzsoc.Division"] <- "Crime.Type"
names(countCrime) <- c("Ethnicity", 'Abduction/Harassment', "Assault", "Endagerment", "Fraud", "Homicide",
                       "Drugs", "Miscellaneous", "Govt/Justice", "Weapons", "Property/Environment", "Public Order",
                       "Robbery/Extortion", "Sexual Assault", "Theft", "Traffic/Vehicle", "Break and Enter")

#Crimes committed by Europeans and Maori

eu.maCrime <- countCrime[c(3,6)]
eu.maCrime <- as.data.frame(t(eu.maCrime))

names(eu.maCrime)[1] <- "European"
names(eu.maCrime)[2] <- "Maori"
eu.maCrime <- eu.maCrime[-1,]
eu.maCrime

#Changing index

eu.maCrime$Type <- rownames(eu.maCrime)
eu.maCrime$Index <- 1:nrow(eu.maCrime)
row.names(eu.maCrime) <- eu.maCrime$Index
eu.maCrime$Index <- NULL

#Pivoting data

eu.maCrime <- eu.maCrime %>% select(Type, Maori, European) %>% 
  pivot_longer(cols = c(Maori, European), names_to = "Var", values_to = "Val")

#Changing datatype of Val from character to integer

eu.maCrime$Val <- as.integer(eu.maCrime$Val)

#Plot

ggplot(data = eu.maCrime, mapping = aes(x=Type, y=Val, fill=Var)) +
  geom_bar(stat = "identity", position=position_dodge()) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Types of Crimes Committed by Maori and Europeans", fill = "Ethnicity") + 
  xlab("Type") + ylab("Frequency")
```

The variables along the x-axis refer to the types of crimes committed by particular ethnicities. The naming of types have been simplified to allow easy fit onto the graph.The y-axis refers to the number of times these types of crimes occurred. 

It is interesting that despite the Maori population being far smaller than the European population in New Zealand, they have similar crime rates, even exceeding Europeans in certain crime types, such as theft and assault. 


Types of crimes committed in certain regions:
```{r, echo=FALSE}
#Types of crimes committed per region

crimeLocation <- OD[ ,c(2,17)]
names(crimeLocation)[1] <- "Crime.Type"
names(crimeLocation)[2] <- "Location"

countCrime.location <- dcast(setDT(crimeLocation, keep.rownames = TRUE), Location ~ Crime.Type, length)
names(countCrime.location) <- c("Location", 'Abduction/Harassment', "Assault", "Endagerment", "Fraud", "Homicide",
                       "Drugs", "Miscellaneous", "Govt/Justice", "Weapons", "Property/Environment", "Public Order",
                       "Robbery/Extortion", "Sexual Assault", "Theft", "Traffic/Vehicle", "Break and Enter")

#Crimes for Aukland and Wellington

countCrime.location <- countCrime.location[c(1,14), ]
countCrime.location <- as.data.frame(t(countCrime.location))
names(countCrime.location)[1] <- "Auckland"
names(countCrime.location)[2] <- "Wellington"
countCrime.location <- countCrime.location[-1,]
countCrime.location

#Changing Index

countCrime.location$Type <- rownames(countCrime.location)
countCrime.location$Index <- 1:nrow(countCrime.location)
row.names(countCrime.location) <- countCrime.location$Index
countCrime.location$Index <- NULL

#Pivoting data

countCrime.location <- countCrime.location %>% select(Type, Auckland, Wellington) %>% 
  pivot_longer(cols = c(Auckland, Wellington), names_to = "Var", values_to = "Val")

#Changing datatype of Val from character to integer

countCrime.location$Val <- as.integer(countCrime.location$Val)

#Plot

ggplot(data = countCrime.location, mapping = aes(x=Type, y=Val, fill=Var)) +
  geom_bar(stat = "identity", position=position_dodge()) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Types of Crimes Committed in the Auckland City and Wellington Districts", fill = "District") + 
  xlab("Type") + ylab("Frequency")
```

It is important to note that Auckland here refers to the Auckland City district, not the entirety of Auckland. Wellington in this instance refers to the entirety of the Wellington district, which includes areas like Upper Hutt. 

Each of the variables along the x-axis are types of crimes that are committed. The naming of types have been simplified to allow easy fit onto the graph. The y-axis is the number of times specific crimes have been committed in the district. 
It is interesting to note that although the Auckland City District has the greatest population out of all possible districts, it has a very similar number of crimes compared to the Wellington District.


```{r, echo=FALSE}
#Age group of offenders

crimeAge <- PS[ ,c(2,6)]
names(crimeAge)[1] <- "Crime.Type"
names(crimeAge)[2] <- "Age"

countCrime.Age <- dcast(setDT(crimeAge, keep.rownames = TRUE), Age ~ Crime.Type, length)
names(countCrime.Age)[names(countCrime)=="Anzsoc.Division"] <- "Crime.Type"

countCrime.Age$Total <- rowSums(countCrime.Age[,2:17])
# countCrime.Age

countCrime.Age <- countCrime.Age[ , c(1, 18)]
countCrime.Age <- countCrime.Age[c(2,3,4,5,7,8,9,11,12,13,14,15,16,17), ]

ggplot(data = countCrime.Age, mapping = aes(x=Age, y=Total)) +
  geom_bar(stat = "identity", fill = "#3366CC") +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Total Crimes Committed by Age Group") + 
  xlab("Age") + ylab("Frequency")
```
Note ages, 0-4, 5-9, NotApplicable and NotSpecified were removed as they were so small compared to the other age groups.

The x-axis refers to the age groups of offenders and the y-axis is the total number of crimes recorded for that specific age group.

A point of interest is that crimes occur the most commonly at ages 15 through to 29, and then begins to taper off as you get older. There is a steady decline in crimes commited as you reach age 30, at a decreasing rate. 


Crimes committed by youths (15-24) and people aged 40-54

```{r, echo=FALSE}
#Types of crimes for youths i.e. 15-24

crimeAge <- PS[ ,c(2,6)]
names(crimeAge)[1] <- "Crime.Type"
names(crimeAge)[2] <- "Age"

countCrime.Age <- dcast(setDT(crimeAge, keep.rownames = TRUE), Age ~ Crime.Type, length)
names(countCrime.Age)[names(countCrime)=="Anzsoc.Division"] <- "Crime.Type"
names(countCrime.Age) <- c("Age", 'Abduction/Harassment', "Assault", "Endagerment", "Fraud", "Homicide",
                                "Drugs", "Miscellaneous", "Govt/Justice", "Weapons", "Property/Environment", "Public Order",
                                "Robbery/Extortion", "Sexual Assault", "Theft", "Traffic/Vehicle", "Break and Enter")

countCrime.Age <- countCrime.Age[c(3,4), ]
countCrime.Age <- as.data.frame(t(countCrime.Age))

#Crimes for specific ages

names(countCrime.Age)[1] <- "15-19"
names(countCrime.Age)[2] <- "20-24"
countCrime.Age <- countCrime.Age[-1,]

#Changing Index

countCrime.Age$Type <- rownames(countCrime.Age)
countCrime.Age$Index <- 1:nrow(countCrime.Age)
row.names(countCrime.Age) <- countCrime.Age$Index
countCrime.Age$Index <- NULL

countCrime.Age$"15-19" <- as.integer(countCrime.Age$"15-19")
countCrime.Age$"20-24" <- as.integer(countCrime.Age$"20-24")

countCrime.Age$Total <- rowSums(countCrime.Age[,c(1,2)])
countCrime.Age <- countCrime.Age[ ,c(3,4)]

ggplot(data = countCrime.Age, mapping = aes(x=Type, y=Total)) +
  geom_bar(stat = "identity", fill = "#3366CC") +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Crimes Committed by Youths (15-24)") + 
  xlab("Type of Crime") + ylab("Frequency")

#----------------------------------------------------------------------------------------

#Crimes for older age

crimeOldAge <- PS[ ,c(2,6)]
names(crimeOldAge)[1] <- "Crime.Type"
names(crimeOldAge)[2] <- "Age"

countCrime.OldAge <- dcast(setDT(crimeOldAge, keep.rownames = TRUE), Age ~ Crime.Type, length)
names(countCrime.OldAge)[names(countCrime)=="Anzsoc.Division"] <- "Crime.Type"
names(countCrime.OldAge) <- c("Age", 'Abduction/Harassment', "Assault", "Endagerment", "Fraud", "Homicide",
                           "Drugs", "Miscellaneous", "Govt/Justice", "Weapons", "Property/Environment", 
                           "Public Order","Robbery/Extortion", "Sexual Assault", "Theft", "Traffic/Vehicle", 
                           "Break and Enter")

countCrime.OldAge <- countCrime.OldAge[c(8,9,10,11), ]
countCrime.OldAge <- as.data.frame(t(countCrime.OldAge))

#Crimes for specific ages

names(countCrime.OldAge)[1] <- "40-44"
names(countCrime.OldAge)[2] <- "45-49"
names(countCrime.OldAge)[3] <- "50-54"
names(countCrime.OldAge)[4] <- "55-59"
countCrime.OldAge <- countCrime.OldAge[-1,]

#Changing Index

countCrime.OldAge$Type <- rownames(countCrime.OldAge)
countCrime.OldAge$Index <- 1:nrow(countCrime.OldAge)
row.names(countCrime.OldAge) <- countCrime.OldAge$Index
countCrime.OldAge$Index <- NULL

countCrime.OldAge$"40-44" <- as.integer(countCrime.OldAge$"40-44")
countCrime.OldAge$"45-49" <- as.integer(countCrime.OldAge$"45-49")
countCrime.OldAge$"50-54" <- as.integer(countCrime.OldAge$"50-54")
countCrime.OldAge$"55-59" <- as.integer(countCrime.OldAge$"55-59")

countCrime.OldAge$Total <- rowSums(countCrime.OldAge[,c(1,2,3,4)])
countCrime.OldAge <- countCrime.OldAge[ ,c(5,6)]

ggplot(data = countCrime.OldAge, mapping = aes(x=Type, y=Total)) +
  geom_bar(stat = "identity", fill = "#CC3300") +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Crimes Committed by people aged 40-54") + 
  xlab("Type of Crime") + ylab("Frequency")

#--------------------------------------------------------------------------------------------

#Comparison of age groups via stacked bar chart

#Renaming total columns

youngCrime <- countCrime.Age
oldCrime <- countCrime.OldAge
names(youngCrime)[2] <- "Youth"
names(oldCrime)[2] <- "Old"

#Combining young and old statistics

compareCrime <- cbind(youngCrime, oldCrime)
compareCrime[3] <- NULL

#Pivot to enable plotting on ggplot

compareCrime.edit <- compareCrime %>% select(Type, Youth, Old) %>% 
  pivot_longer(cols = c(Youth, Old), names_to = "Var", values_to = "Val")

ggplot(data = compareCrime.edit, mapping = aes(x=Type, y=Val, fill = Var)) +
  geom_bar(stat = "identity", position = "fill") +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Proportion of Crimes Committed by Youths (15-24) and Old (40-59)", fill = "Age group") + 
  xlab("Type of Crime") + ylab("Proportion")
```

Each of the variables along the x-axis are types of crimes that are committed. The naming of types have been simplified to allow easy fit onto the graph. The y-axis is the number of times specific crimes have been committed by particular age brackets.

The most common types of crime for youths to commit are assault, theft, and traffic/vehicle offences, which are similar to the most common crimes committed by people in the age bracket 40-54, with the exception of Offences Against Justice Procedures, Govt Sec and Govt Ops (Govt/Justice). Theft is also much more common for youths when compared to the older age group.

Assault is the most common crime for the older age bracket, whereas traffic/vehicle offences are the most common for youths. 

# Police Data (Victimisations Time and Place):


```{r, echo=FALSE}
policedata<-read.csv(file = "Policedata.csv")

#str(policedata)
summary(policedata)
sum(is.na(policedata))

```


```{r, echo=FALSE, include=FALSE}

#Count of different type of crimes

counts <- table(policedata$ANZSOC.Division)
barplot(counts, main="Count of Different type of crimes",
  xlab="Crime type", names.arg=c("Abduction/Harassment","Acts Cause Injury","Robbery/Extortion","Sexual Assault","Theft","Unlawful Entry:Break/Enter"))

#The variables along the x-axis refer to the types of crimes committed. The naming of types have been simplified to allow easy fit onto the #graph. The y-axis refers to the count of times these types of crimes occurred. Theft is the most occurred types of crime committed followed by #Unlawful Entry:Break/Enter

```


```{r, echo=FALSE, include=FALSE}
#Count of occurrence of Territorial Authorities (area of the crime committed)

counts <- table(policedata$Territorial.Authority)
barplot(counts, main="Count of occurrence of Territorial Authorities",
  xlab="Territorial Authorities", names.arg=c("Carterton","Upper Hutt","Lower Hutt","Kapiti Coast ","Porirua ","South Wairarapa ","Tararua ","Masterton ","Wellington "))

```


```{r, echo=FALSE, include=FALSE}
#Types of Crimes that occurred all around Wellington

counts <- table(policedata$ANZSOC.Division, policedata$Territorial.Authority)
barplot(counts, main="Crime type by Territorial authority",
xlab="Territorial Authority", col = c("red","blue","orange","green","purple","yellow"),
names.arg=c("Carterton","Upper Hutt","Lower Hutt","Kapiti Coast ","Porirua ","South Wairarapa ","Tararua ","Masterton ","Wellington "),
legend = c("Abduction/Harassment","Acts Cause Injury","Robbery/Extortion","Sexual Assault","Theft","Unlawful Entry:Break/Enter"))

#Each of the variables along the x-axis are area of the crime committed. The naming of types have been simplified to allow easy fit onto the #graph. The y-axis is the the count of times these types of crimes occurred in this areas. It seems like there aren't any links or patterns #between the type of the crime and the area of the crime committed. Most number of all different types of crimes being committed is in #wellington followed by next by Lower Hutt district.

```



# Ethical Considerations
1.) Many ethical remarks can be raised about the morality of undertaking feature analysis on perpetrators of different divisions of crime. 

 + One of the most pronounced issues would be applying learnings from patterns in data to individuals in the real world. Putting firm labels on features that offenders appear to have in common is inherently discriminatory. Just as equally, we may be able to identify trends (feature patterns, or offender commonalities) that might shift focus more onto pre-crime interventions (for subgroups discovered to be "more likely" to tend towards crime), and less onto post-crime rehabilitation.
 
 + An additional ethical consideration is the moral quandary of choosing whether or not to use offender data to predict future trends in crime. A positive reason for doing so might be to reduce victimisations for certain subgroups of people. However, a significant consequence of this would be the influence of data-driven bias leading to selffulfilling scenarios. The example here would be: An upwards trend in crime is identified for (insert police district). More police are tasked with community surveillance. More surveillance leads to more detected crimes for said district, and fewer for another (because police are a finite resource). This district becomes more marginalised, causing the members of its community to fall into a lower-income decile.  

# Privacy Considerations  

2.) Privacy issues that may arise when using these datasets:

 + Ensure that suppressed information cannot be related back to people it is pertaining to.  
 + Data collection and use must meet the standards listed in the NZ Privacy Act 1995, and the Statistics Act of 1975 – amended 2012  

# Security Considerations

3.) Steps that could be taken to keep the project data and results secure:

 + Using complex passwords provide essential protection from fraud and hackers from accessing our data as guessing passwords is one of the most common method hackers use to breaks into computers.  
 + Data encryption will further secure the data by encoding the data so that it remains hidden from unauthorized users. Encryption protects private and sensitive information and enhances the security of communication between clients and servers.   
 + Allowing permitted individuals only to access those computer areas, e.g. by restricting entry by ID cards or magnetic swipe cards to these areas  

# Individual Contributions
1.) Maria DaRocha:   
+ Background Description Victimisations (Police Stations) & Unique Victims (Police Stations)  
+ EDA Victimisations & Unique Victims (Police Stations)  
+ Ethical considerations  
+ Code integration  
+ Report consolidation  
+ Project Organisation (Trello)  
+ Version control (GitHub)
  
2.) George Pope:   
+ Background Description Offender Proceedings (Offender Demographics) and Offender Proceedings (Police Stations)  
+ EDA Offender Proceedings (Offender Demographics) and Offender Proceedings (Police Stations)  
+ Privacy considerations   
+ References   
+ Version control (GitHub) 
  
3.) Usanthini Sasikumar:  
+ Background Description Police Data (Victimisation Time and Place)  
+ EDA Police Data (Victimisation Time and Place)  
+ Security Considerations  

# References

 + NZ Police. (2017). Victimisations Time and Place (V2:00) [Datafile]. Available from NZ Police Website: https://www.police.govt.nz/about-us/publications-statistics/data-and-statistics/policedatanz/victimisation-time-and-place
 + NZ Police. (2017). Offender Proceedings (Police Stations). [Datafile]. Available from NZ Police Website: https://www.police.govt.nz/about-us/statistics-and-publications/data-and-statistics/proceedings-police-stations
 + NZ Police. (2017). Offender Proceedings (offender demographics). [Datafile]. Available from NZ Police Website: https://www.police.govt.nz/about-us/publications-statistics/data-and-statistics/policedatanz/proceedings-offender-demographics
 + NZ Police. (2017). Unique Victims (demographics) (02.00) [Datafile]. Available from NZ Police Website: https://www.police.govt.nz/about-us/publications-statistics/data-and-statistics/policedatanz/unique-victims-demographics
 + NZ Police. (2017). Unique Offenders (demographics) [Datafile]. Available from NZ Police Website: https://www.police.govt.nz/about-us/publications-statistics/data-and-statistics/policedatanz/unique-offenders-demographics
 + NZ Police. (2017). Victimisations (Police Stations) [Datafile]. Available from NZ Police Website: https://www.police.govt.nz/about-us/statistics-and-publications/data-and-statistics/victimisations-police-stations
 + NZ Police. (2017). Unique Victims (Police Stations) (01.00). [Datafile]. Available from NZ Police Website: https://www.police.govt.nz/about-us/statistics-and-publications/data-and-statistics/unique-victims-police-stations
 + NZ Police. (2020). Covid-19 Response Data (01.01) [Datafile]. Available from NZ Police Website: https://www.police.govt.nz/about-us/statistics-and-publications/data-and-statistics/covid-19-response


