---
title: "Recorded Crime Victim Statistics (RCVS), Police Stations"
subtitle: "(Datasets Policedata.nz: AEG_Full_Data_data.csv, UAEG_Full_Data_data.csv)"
author: "EDA, Maria DaRocha"
date: "8/30/2020"
output: pdf_document
---

Among the datasets we decided to work on, two of them were: Victimisations (Police Stations), and Unique Victims (Police Stations). These datasets were of particular interest because oftentimes when we discuss crime, the majority of our conversations revolve around the offender: their background, their demographic, etcetera. While these details may help us better understand an individual’s propensity towards crime, focusing solely on the offender can mean overlooking the other involved party: their victims. Thoughtfully done, analyzing information about victims can lead to a greater understanding of commonalities between victims of similar crimes. This layer of insight could help us better understand what might make someone more vulnerable to crime, or susceptible to being targeted by a specific type of crime. The datasets themselves consisted of numerical and categorical data. The data is best suited to time series modelling, as it is centered around victimisations and unique victims within a specific time period (2015-2020).  

While there were no missing values in the Victimisations (Police Stations) dataset, there were some corrupted entries in which a categorical variable (Ages 10-19) had become stored as a date (Oct-19). This issue could be related to the files being exported from policedata.nz as ‘Microsoft Excel csv files,’ instead of regular csv files. Data integration took place on an as-needed basis. For example, when comparing ethnicity and crime division features in each dataset, the four features were merged into a single dataframe. The Unique Victims dataset was missing certain values.

```{r,echo=FALSE,message=FALSE,warning=FALSE}
require(knitr)
require(formatR)
# Set so that long lines in R will be wrapped:
opts_chunk$set(tidy.opts=list(width.cutoff=40),tidy=TRUE)
opts_knit$set(global.par = TRUE)
## Reduce the amount of whitespace above each plot
if(is_latex_output()) {
  plot_default <- knit_hooks$get("plot")
  knit_hooks$set(plot = function(x, options) { 
    x <- c("\\vspace{-2mm}", plot_default(x, options))
  })
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
VD <- read.csv("AEG_Full_Data_data.csv")
UVD <- read.csv("UAEG_Full_Data_data.csv")

library(ggplot2)
library(dplyr)
library(tidyr)
library(data.table)
library(RColorBrewer)
```

# General EDA, Victimisations (Police Stations): 
## Missing Values, Data Types, Etc.

```{r}
# Trivial investigation
summary(VD)
test_NA_VD <- sum(is.na(VD))
test_NA_UVD <- sum(is.na(UVD))

test_NA_UVD
test_NA_VD

# Not expressive of corrupt data (age values, VD)
```

  
# Victimisations (Police Stations): 
## Crime Division, Ethnicities of Victims   

In the victimisations (by police stations) dataset, we chose to explore ethnicity’s connection to each crime division. If patterns are found in later, deeper analysis - then we will be able to conclude some correlation between ethnicity and a particular type of crime. In a real-world application, this would help identify a population that may be more vulnerable to certain crimes. Initially, we explored the data using the classifications for ethnicity provided.
  
```{r}
division_ethnicity <- VD[ , c(2, 9)]
names(division_ethnicity)[names(division_ethnicity)=="Anzsoc.Division"] <- "Division"
names(division_ethnicity)[names(division_ethnicity)=="Abduction, Harassment and Other Related Offences Against a Person"] <- "Abduction"

victimisation_count <- dcast(setDT(division_ethnicity, keep.rownames = TRUE), Ethnicity ~ Division, length)

kable(head(victimisation_count))
```


```{r}
# Visual Observation: victimisation ONE crime division with Bar Chart
ggplot(data = victimisation_count, mapping = aes(x=victimisation_count$Ethnicity, y=victimisation_count$`Abduction`, fill=victimisation_count$Ethnicity)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle =70, hjust=1)) +
  labs(title = "Abduction and Harassment, Victimisations", subtitle = '*Includes Related Offences (2015-2020)', fill = "Ethnicity") + xlab("Ethnicity") + ylab("Recorded Victimisations")
```
```{r}
# Visual Observation: victimisation ONE crime division with Bar Chart
ggplot(data = victimisation_count, mapping = aes(x=victimisation_count$Ethnicity, y=victimisation_count$`Sexual Assault and Related Offences`, fill=victimisation_count$Ethnicity)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle =70, hjust=1)) +
  labs(title = "Sexual Assault, Victimisations", subtitle = '*Includes Related Offences (2015-2020)', fill = "Ethnicity") + xlab("Ethnicity") + ylab("Recorded Victimisations")
```
  
# Victimisations (Police Stations): 
## Crime Division, (Aggregated) Ethnicities of Victims   

Upon seeing the data heavily reflect European, Maori, and unspecified (or suppressed) ethnicities, it seemed appropriate to aggregate non-Maori and non-Eurpoean ethnic groups into a "Minority (Other)" category. This aggregation provides a visual model which improves the comparability of minority groups that are less prevalent New Zealand. In doing this, we ensure that individuals who separately fall into the (now aggregated) "Minority (other)" category are fairly represented - and that visually, the model no longer reflects any one minority category appearing less targeted than Maori or European persons.

```{r}
# Aggregated Minority (Other)
# Retain 'Not Stated', 'Not Applicable', 'Not Elsewhere Classified'
# 'Organisation', 'Suppressed'
aggregation <- c(0,0,1,0,0,2,0,3,4,5,6,0,7)
victimisation_count_copy <- victimisation_count
victimisation_count_copy$Ethnicity <- aggregation
aggregate_victimisation <- aggregate(x=victimisation_count_copy, by=list(victimisation_count_copy$Ethnicity), FUN = sum)
names(aggregate_victimisation)[names(aggregate_victimisation)=="Abduction, Harassment and Other Related Offences Against a Person"] <- "Abduction"
aggregate_victimisation$Ethnicity <- c("Minority (Other)", "European", "Maori", "Not Applicable", "Not Elsewhere Classified", "Not Stated", "Organisation", "Suppressed")
# drop added grouping
aggregate_victimisation <- aggregate_victimisation[,-(1)]
kable(head(aggregate_victimisation))
```


```{r}
# Visual Observation of ONE division with Bar Chart
ggplot(data = aggregate_victimisation, mapping = aes(x=Ethnicity, y=`Abduction`, fill=Ethnicity)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 75, hjust=1)) +
  labs(title = "Abduction and Harassment, Victimisations", subtitle = '*Includes Related Offences (2015-2020)', fill = "Ethnicity") + xlab("Ethnicity") + ylab("Recorded Victimisations")
```
  
# Victimisation (Police Stations): 
## Crime Division to Year of Report  

The next interesting feature of the data is "Year.Month" - i.e. the basis for our time series. In future analysis, we will investigate this feature in greater detail. For now, we can aggregate these values to just the year on record and observe basic trends.

```{r}
victimisation_series <- VD[ , c(2, 5)]

names(victimisation_series)[names(victimisation_series)=="Anzsoc.Division"] <- "Division"
series_count <- dcast(setDT(victimisation_series, keep.rownames = TRUE), Year.Month ~ Division, length)

```

```{r}
agg_series_count <- series_count
years <- c(2015, 2016, 2017, 2018, 2019, 2020)
for (k in years){
  year <- grepl(k, agg_series_count$Year.Month)
  index <- 0
  for (y in year){
    index <- index + 1
    if (y==TRUE){
      agg_series_count$Year.Month[index] = k
    }
  }
}  

names(agg_series_count)[names(agg_series_count)=="Year.Month"] <- "Year"
names(agg_series_count)[names(agg_series_count)=="Abduction, Harassment and Other Related Offences Against a Person"] <- "Abduction"

agg_series_count <- aggregate(cbind(`Abduction`, `Acts Intended to Cause Injury`, `Robbery, Extortion and Related Offences`, `Sexual Assault and Related Offences`, `Theft and Related Offences`, `Unlawful Entry With Intent/Burglary, Break and Enter`) ~ Year, agg_series_count, sum)

kable(head(agg_series_count))
```


```{r}
# Set up visualisation
all_colnames <- c("Date","Abduction","Intended_Injury","Robbery","Sexual_Assault","Theft","Burglary")
index <- 0
for (col in all_colnames){
  index <- index + 1
  names(agg_series_count)[index] <- all_colnames[index]
}

agg_series_count <- agg_series_count %>%
  select(Date,Abduction,Intended_Injury,Robbery,Sexual_Assault,Theft,Burglary) %>%
  gather(key = variable, value = value, -Date)

```
  
# Time Series, Victimisations (Police Stations):
## Change in Recorded Victimisations by Crime Division (Annually) 
  
```{r}
# Draw Time Series
ggplot(agg_series_count, aes(x = Date, y = value, group = variable)) + 
  geom_line(aes(color = variable, linetype = variable)) + 
  scale_color_brewer(palette = "Dark2") +
  theme(axis.text.x=element_text(angle = 90, size = 6, hjust=1)) +
  labs(title = "Victimisations by Crime Division by Year", subtitle = 'Time Series Data (2015-2020)', fill = "Crime Division") + xlab("Year") + ylab("Recorded Victimisations")

```

# Unique Victims (Police Stations): 
## Crime Division, (Aggregated) Ethnicities of Victims  
  

```{r}
#head(UVD)
uv_count <- UVD[ , c(2, 8)]
names(uv_count)[names(uv_count)=="Anzsoc.Division"] <- "Division"

uv_count <- dcast(setDT(uv_count, keep.rownames = TRUE), Ethnicity ~ Division, length)

# Aggregated Minority (Other)
# Retain 'Not Stated'
aggregation <- c(0,0,1,0,0,2,0,3,4,5,6,0,7)
uv_count_copy <- uv_count
uv_count_copy$Ethnicity <- aggregation
aggregate_uv <- aggregate(x=uv_count_copy, by=list(uv_count_copy$Ethnicity), FUN = sum)
names(aggregate_uv)[names(aggregate_uv)=="Abduction, Harassment and Other Related Offences Against a Person"] <- "Abduction"
aggregate_uv$Ethnicity <- c("Minority (Other)", "European", "Maori", "Not Applicable", "Not Elsewhere Classified", "Not Stated", "Organisation", "Suppressed")
# drop added grouping
aggregate_uv <- aggregate_uv[,-(1)]

kable(head(aggregate_uv))
```
```{r}
# Visual Observation of ONE division with Bar Chart
ggplot(data = aggregate_uv, mapping = aes(x=Ethnicity, y=`Abduction`, fill=Ethnicity)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 75, hjust=1)) +
  labs(title = "Abduction and Harassment, Unique Victims", subtitle = '*Includes Related Offences (2015-2020)', fill = "Ethnicity") + xlab("Ethnicity") + ylab("Recorded Victimisations")

```

# Unique Victims & Victimisations (Police Stations):
## Distributional Analysis with Standard Deviation

For the final part of the exploratory analysis on this data, it seemed like a reasonable inquiry to test the comparablity of the victimisations and unique victims datasets (in regards to variability). Ethnicity was used as a dummy variable to track the consistency of a single feature's variation across the two datasets. We would expect, if the datasets are comparable, for the standard deviations to exhibit the same (or relatively similar trends). This proved to be the case - and going forward, it will be possible to use this test holding other features of interest constant, and potentially use the two datasets interchangeably. **This is invaluable knowledge for avoiding temporally and computationally expensive (unnecessarily redundant) investigations.**

```{r}
distribution_names <- colnames(aggregate_uv[1,])

distribution_names <- distribution_names[-1]


# Comparing Distributions
set.seed(1234)
for (n in 1:length(distribution_names)){
  name <- distribution_names[n]
  
  uv_vals <- aggregate_uv[ , grepl(name, names(aggregate_uv))]
  v_vals <- aggregate_victimisation[ , grepl(name, names(aggregate_uv))]
  
  uv_group <- data.frame(x = aggregate_uv$Ethnicity, 
                  y = uv_vals)
  uv_group$group=c(0,0,0,0)
  uv_group <- uv_group %>% mutate_at(c("y"), ~(scale(.) %>% as.vector))
  
  v_group <- data.frame(x = aggregate_victimisation$Ethnicity, 
                    y = v_vals)
  v_group$group=c(1,1,1,1)
  v_group <- v_group %>% mutate_at(c("y"), ~(scale(.) %>% as.vector))
  
  # Merge dataframes
  distribution_analysis <- merge(uv_group, v_group, by.x="x", by.y="x")

  # Pivot data
  distribution_analysis_copy <- distribution_analysis %>% select(x, y.x,y.y) %>% 
  pivot_longer(cols = c(y.x,y.y), names_to = "variable", values_to = "value")
  # print(distribution_analysis_copy)
  
  # Plot
  p <- ggplot(distribution_analysis_copy, aes(x = x, y = value, group = variable)) + 
  geom_line(aes(color = variable, linetype = variable)) + 
  scale_color_brewer(palette = "Dark2") +
  theme(axis.text.x=element_text(angle = 75, size = 10, hjust=1), legend.position = "none") +
  labs(title = name, 
       subtitle = 'Fluxuation in Distribution of Ethnicities: \nUnique Victims and Victimisations (2015-2020)', fill = name) + xlab("Ethnicity") + ylab("Stnd Deviation (Mean)")
  show(p)
}

```




